---
title: "POL_279_Assign4"
author: "E Zufall"
date: "2/22/2023"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(statnet)
```

## Individual Network Characteristics

Density
The density of a binary network is given by $\frac{\sum_{i=1}^{n-1} \sum_{j=i+1}^{n} S_{ij}}{n(n-1)}$

The density is minimized (0) when there are no edges in the network, as follows.
```{r density_low}
adj_mat1 <- matrix(rep(0,100), nrow=10)
network1 <- network(adj_mat1)
print(network.density(network1))
```

The density is maximized (1) when every possible edge is drawn, as follows.
```{r density_high}
adj_matn <- matrix(rep(1,100), nrow=10)
networkn <- network(adj_matn)
print(network.density(networkn))
```


Global Transitivity
The global transitivity of a binary network is given by 
$\frac{6\sum_{i\neq j\neq k} t_{ijk}}{(n-1)(n-2)}$ where $t_{ijk} = s_is_js_k$

Components
Components are found the reachability matrix
$R = \sum_{i=1}^{n-1}S^i$, such that for component $q$, matrix entry $r_{ij} \neq 0$ for all nodes $i$, $j \in q$, and 0 for all $i \in q, j\notin q$, or $i \notin q, j\in q$. 

Polarization

Group (Degree) Centralization

## Network Characteristic Pairs
Density and Transitivity


Density and Number of Components
Transitivity and Number of Components
Density and Polarization
Transitivity and Polarization
Polarization and Group (Degree) Centralization

```{r density_transitivity}
adj_mat1 <- matrix(rep(0,100), nrow=10)
network1 <- network(adj_mat1)
network.density(network1)
gtrans(network1)

adj_mat2 <- matrix(c(c(0,1,0,0,0,0,0,0,0,0),
                     c(0,0,1,0,0,0,0,0,0,0),
                     c(0,0,0,1,0,0,0,0,0,0),
                     c(0,0,0,0,1,0,0,0,0,0),
                     c(0,0,0,0,0,1,0,0,0,0),
                     c(0,0,0,0,0,0,1,0,0,0),
                     c(0,0,0,0,0,0,0,1,0,0),
                     c(0,0,0,0,0,0,0,0,1,0),
                     c(0,0,0,0,0,0,0,0,0,1),
                     c(1,0,0,0,0,0,0,0,0,0)), byrow=T, nrow=10)
network2 <- network(adj_mat2)
network.density(network2)
gtrans(network2)
plot(network2)

adj_mat3 <- matrix(c(c(0,1,1,0,0,0,0,0,0,0),
                     c(1,0,1,0,0,0,0,0,0,0),
                     c(1,1,0,0,0,0,0,0,0,0),
                     c(0,0,0,0,1,1,0,0,0,0),
                     c(0,0,0,1,0,1,0,0,0,0),
                     c(0,0,0,1,1,0,0,0,0,0),
                     c(0,0,0,0,0,0,0,1,1,0),
                     c(0,0,0,0,0,0,1,0,1,0),
                     c(0,0,0,0,0,0,1,1,0,0),
                     c(0,0,0,0,0,0,0,0,0,0)), byrow=T, nrow=10)
network3 <- network(adj_mat3)
network.density(network3)
gtrans(network3)
plot(network3)

adj_matn <- matrix(rep(1,100), nrow=10)
networkn <- network(adj_matn)
network.density(networkn)
gtrans(networkn)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
